Open Clusters Characterization in Gaia DR2 Using ML Algorithms

### 2020-12-07 thesis/thesis.tex
thesis: Update Method chapter

\usepackage{apacite}
\usepackage{caption}
\usepackage{csquotes}
\usepackage{eso-pic}
\usepackage{float}
\usepackage{graphicx}
\usepackage{minted}
\usepackage{picture}
\usepackage{subcaption}
\usepackage[nottoc]{tocbibind}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.9\textwidth]{../figures/clusterix/selection.pdf}
  \caption{Typical Clusterix 2.0 control panel.}
  \label{fig:clusterix_control_panel}
\end{figure}

@@

\chapter{Method}

@@

\section{Data Mining}

@@

This download is not limited to the size registered in the catalogue. Instead, a wider region is downloaded for each cluster to include
stars outside the cluster. The idea is that the unsupervised model must be able to clusterize the stars that belongs to the OC and to discard
those outsiders when characterizing the open cluster.

@@

@@

Finally, Figure \ref{fig:hr_diagram_comparison} shows a comparison between the H-R diagram obtained by DEC clustering and the one obtained
using the Clusterix + TOPCAT method.

>>>>

\usepackage{apacite}
\usepackage{caption}
\usepackage{csquotes}
\usepackage{eso-pic}
\usepackage{float}
\usepackage{graphicx}
\usepackage{minted}
\usepackage{picture}
\usepackage{subcaption}
\usepackage{tablefootnote}
\usepackage[nottoc]{tocbibind}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.9\textwidth]{../figures/clusterix/selection.pdf}
  \caption{Typical Clusterix 2.0 region selection panel.}
  \label{fig:clusterix_control_panel}
\end{figure}

@@

\chapter{Method}
\label{chap:method}

@@

\section{Data Mining}
\label{sec:data_mining}

@@

This download is not limited to the size registered in the catalogue.
Instead, a wider region (1.5 times the size of the cluster) is downloaded for each cluster to include stars outside the cluster.
The idea is that the unsupervised model must be able to clusterize the stars that belongs to the OC and to discard
those outsiders when characterizing the open cluster.

@@

Figure \ref{fig:hr_diagram_comparison} shows a comparison between the H-R diagram obtained by DEC clustering and the one obtained
using the Clusterix + TOPCAT method.

@@

Finally, we can refine this selection by filtering those stars which are bellow and above the 0.25 and 0.75 quantiles for each group, respectively.
That way we remove the most doubtful values from the selection.
Figure \ref{fig:melotte_22_filtered} shows the result after filtering DEC clustering selection.

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_filtered_melotte_22.png}
      \caption{H-R diagram}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_filtered_melotte_22.png}
      \caption{Parallax}
    \end{subfigure}
  \end{subfigure}
  \caption{Melotte 22 H-R diagram and parallax histogram with stars outside 0.25 and 0.75 quantiles filtered.}
  \label{fig:melotte_22_filtered}
\end{figure}

Table \ref{tab:results_melotte_22} shows a summary of Melotte 22 with values taken from different sources and methods: SIMBAD Astronomical Database - CDS (Strasbourg),
Clusterix+TOPCAT method, and the results obtained for each step of this method (KMeans, DEC and DEC filtered).

% TODO: Fix table if figures changes
\begin{table}[h!]
  \begin{center}
    \begin{tabular}{l|c|c|c|c}
      \textbf{Source / method} & \textbf{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \textbf{$\mu_{\delta}$ $(mas/yr^{-1})$} & \textbf{$\varpi$ $(mas)$} & \emph{\# stars} \\
      \hline
      Simbad.u-strasbg.fr\tablefootnote{Results have been taken from \protect\citeA[p.~25. Table A.3. Pleiades]{babusiaux2018gaia}} &
        19.997 $\pm$ 0.127 & -45.548 $\pm$ 0.101 & 7.364 $\pm$ 0.005 & 1326 \\
      Clusterix+TOPCAT & 19.98 $\pm$ 1.25 & -45.47 $\pm$ 1.48 & 7.33 $\pm$ 0.21 & 168 \\
      KMeans & 20.25 $\pm$ 35.16 & -38.01 $\pm$ 40.16 & 7.23 $\pm$ 2.35 & 1378 \\
      DEC & 23.67 $\pm$ 38.33 & -46.23 $\pm$ 44.54 & 8.04 $\pm$ 2.59 & 878 \\
      DEC (filtered) & 19.50 $\pm$ 8.59 & -44.23 $\pm$ 8.17 & 7.42 $\pm$ 0.1 & 438 \\
    \end{tabular}
    \caption{Melotte 22. The parameters shown are proper motion in right ascension and declination, parallax and number of stars with their respective deviations.}
    \label{tab:results_melotte_22}
  \end{center}
\end{table}
